---
title: "NPFMC SAFE catalog audit"
format:
  html:
    toc: true
    number-sections: true
    lightbox: true
    code-fold: true
    code-summary: "Show R code"
execute:
  echo: true
  warning: false
  message: false
---

```{r}
source("npfmcSAFE.R")
library(dplyr)
library(gt)
library(scales)
```

## Purpose

This document builds a combined catalog from:

- the NPFMC SAFE reports page
- the NOAA Fisheries North Pacific groundfish assessment index
- the linked NOAA annual groundfish pages

It is designed to audit naming consistency, link structure, and year anomalies, with special attention to the unusual 2025 groundfish page.

## Build catalog

```{r}
catalog_path <- file.path("outputs", "safe_catalog.csv")

if (file.exists(catalog_path)) {
  safe_catalog <- readr::read_csv(catalog_path, show_col_types = FALSE)
} else {
  safe_catalog <- build_safe_catalog(crawl_noaa = TRUE, quietly = TRUE)
  write_safe_outputs(safe_catalog, out_dir = "outputs")
}

safe_summaries <- summarise_safe_catalog(safe_catalog)
```

## Headline counts

These summary counts show the current size of the catalog, how many links were flagged as unusual, and how much temporal and stock coverage the combined crawl currently reaches.

```{r}
headline <- tibble::tibble(
  metric = c(
    "Total unique linked documents",
    "Anomalous links flagged",
    "Distinct stocks inferred",
    "Distinct effective years inferred"
  ),
  value = c(
    nrow(safe_catalog),
    sum(safe_catalog$is_anomalous, na.rm = TRUE),
    dplyr::n_distinct(stats::na.omit(safe_catalog$stock)),
    dplyr::n_distinct(stats::na.omit(safe_catalog$true_year))
  )
)

headline |>
  gt() |>
  fmt_number(columns = value, decimals = 0)
```

## Documents by year

This plot gives the fastest read on whether the crawl is reaching the expected historical range and whether any NOAA entry-point changes created visible gaps in year coverage.

```{r}
plot_safe_docs_by_year(safe_catalog)
```

## Stock coverage

These views group stocks by FMP first, then by the most relevant sub-area within that FMP. GOA entries are separated into `GOA` versus `Both` where needed, while BSAI entries are separated into `EBS`, `AI`, or `Both`.

```{r}
plot_stock_coverage(safe_catalog, n = Inf, fmp_filter = "BSAI")
```

```{r}
plot_stock_coverage(safe_catalog, n = Inf, fmp_filter = "GOA")
```

## Anomalies

The anomaly table focuses on links where the visible year, directory year, or reused filename pattern does not line up cleanly with the inferred effective year.

```{r}
safe_summaries$anomalies |>
  select(source, section, label, filename, source_year_dir, true_year, anomaly_flags, url) |>
  gt() |>
  fmt_missing(everything(), missing_text = "") |>
  tab_header(title = "Flagged anomalies")
```

## Filename collisions across effective years

Repeated filenames across multiple effective years are worth tracking because they can hide silent document replacement or directory reuse.

For this audit, the effective year is based on the assessment folder or page context, not on any file timestamp. For example, `2003/EBSpollock.pdf` is treated as the 2003 assessment because of the folder name, regardless of the file's later modification date.

The summary table below is meant to be read one filename at a time. Each row says that the same basename appears in multiple assessment years, shows the total number of distinct years, and lists those years explicitly. For example, if `AIpcod.pdf` has `8` years listed, that means the same filename was reused in eight separate assessment-year contexts.

```{r}
filename_collision_summary <- safe_summaries$filename_collisions |>
  group_by(source, filename, n_true_years) |>
  summarise(
    first_year = min(true_year, na.rm = TRUE),
    last_year = max(true_year, na.rm = TRUE),
    years = paste(sort(unique(true_year)), collapse = ", "),
    .groups = "drop"
  ) |>
  arrange(source, filename)

filename_collision_summary |>
  gt() |>
  cols_label(
    n_true_years = "distinct_years",
    first_year = "first_year",
    last_year = "last_year",
    years = "year_list"
  ) |>
  fmt_missing(everything(), missing_text = "")
```

## Stock inventory

This inventory is a compact reference for what the crawler currently infers about stock, area, and section coverage across the combined catalog.

```{r}
safe_summaries$stock_inventory |>
  arrange(section, stock, area) |>
  gt() |>
  fmt_number(columns = n, decimals = 0)
```

## Example: BSAI pollock access history

Pollock is a useful regression test because the file naming has changed over time and the same stock can appear through multiple NOAA and NPFMC entry points.

```{r}
pollock_example <- safe_catalog |>
  filter(
    grepl("Pollock", stock %||% "") | grepl("pollock", label, ignore.case = TRUE),
    grepl("BSAI|Bering", area %||% "", ignore.case = TRUE) |
      grepl("BSAI", label, ignore.case = TRUE)
  ) |>
  mutate(
    access_note = case_when(
      filename == "BSAIpollock.pdf" ~ "Current naming uses BSAIpollock.pdf rather than EBSpollock.pdf",
      filename == "EBSpollock.pdf" ~ "Alternate or older filename found",
      TRUE ~ ""
    )
  ) |>
  select(source, true_year, label, filename, access_note, url) |>
  arrange(desc(true_year), source, label)

pollock_example |>
  gt() |>
  fmt_missing(everything(), missing_text = "")
```

## Example: GOA Pacific cod access history

GOA Pacific cod is a second targeted check because the current NPFMC page mixes ordinary 2025 material with visible 2026 update files under the 2025 directory.

```{r}
pcod_example <- safe_catalog |>
  filter(
    grepl("Pacific cod", stock %||% "", ignore.case = TRUE) |
      grepl("pcod", filename %||% "", ignore.case = TRUE) |
      grepl("Pacific cod", label, ignore.case = TRUE),
    grepl("GOA", area %||% "", ignore.case = TRUE) |
      grepl("GOA", label, ignore.case = TRUE)
  ) |>
  mutate(
    access_note = case_when(
      filename == "GOApcod.pdf" & grepl("\\b2026\\b", label) ~
        "2026 GOA Pacific cod stored under /SAFE/2025/GOApcod.pdf",
      filename == "GOApcod25.pdf" ~
        "Explicit 2025 GOA Pacific cod filename",
      TRUE ~ ""
    )
  ) |>
  select(source, true_year, source_year_dir, label, filename, access_note, url) |>
  arrange(desc(true_year), source, label)

pcod_example |>
  gt() |>
  fmt_missing(everything(), missing_text = "")
```

## Resolved NOAA terminal links

This long table lists terminal assessment links that point to the actual downloadable files rather than wrapper pages. It is the main place to confirm that NOAA `resource/data` cover pages and the REFM historic archive are being converted into live end-document URLs.

```{r}
resolved_noaa_links <- safe_catalog |>
  filter(
    source %in% c("NOAA", "NOAA_REFM"),
    !grepl("fisheries.noaa.gov/resource/data", url, fixed = TRUE)
  ) |>
  arrange(desc(true_year), filename, label) |>
  mutate(
    live_link = sprintf(
      '<a href="%s" target="_blank" rel="noopener noreferrer">open</a>',
      url
    )
  ) |>
  select(source, true_year, filename, live_link, label, stock)

DT::datatable(
  resolved_noaa_links,
  filter = "top",
  escape = FALSE,
  rownames = FALSE,
  class = "compact stripe hover",
  colnames = c("source", "true_year", "filename", "live_link", "label", "stock"),
  options = list(
    pageLength = 25,
    autoWidth = TRUE,
    order = list(list(1, "desc"), list(2, "asc"), list(4, "asc"))
  ),
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: 600;",
    "Resolved NOAA and REFM terminal assessment links"
  )
) |>
  DT::formatStyle(columns = names(resolved_noaa_links), `white-space` = "nowrap")
```

## Notes

- The 2025 NPFMC groundfish page explicitly says the GOA and BSAI groundfish SAFE reports were **not compiled** that year and that catch reports were prepared instead.
- The same page also includes **2026 GOA Pacific cod** material inside the 2025 directory structure, which is why the anomaly checks treat visible link text as authoritative in that case.
- NOAA’s groundfish index page acts as a second entry point by linking yearly pages for 2013–2024, while REFM year pages are crawled directly for 2006–2012 and the REFM historic archive extends coverage farther back.
- NOAA `resource/data` wrappers are dereferenced so the catalog keeps the underlying downloadable assessment files when those links are exposed on the wrapper page.
